Automated iterative forecasting for the Portal Project

### 2018-02-04 manuscript/portal_iterative_forecasting_ms.Rmd

Forecasting the future state of ecological systems is important for management, conservation, and evaluation of our basic understanding of how ecology operates [@clark2001; @tallis2006; @diaz2015; @dietze2017]. In 2001, Clark et al. [-clark2001] called for a more central role of forecasting in ecology, and since then an increasing number of ecological forecasts are being published. However, most of these forecasts are made once, published, and never assessed or updated. Without assessment, we have limited information on how much confidence to place in our predictions; without regular updates, forecasts lack the most up-to-date information as conditions change [@dietze2018]. This lack of both regular assessment and active updating has limited the progress of ecological forecasting and hindered our ability to make useful and reliable predictions. More regular updating and assessment will advance ecological forecasting as a field by speeding up the identification of the best models for forecasting specific issues in specific ecosystems and contribute overall to our understanding of how to best design forecasting approaches for ecology. For ecological forecasting to mature as a field, we need to change how we produce and interact with forecasts, creating a more dynamic interplay between model development, prediction generation, and incorporation of new data and information [@dietze2018].

With the goal of making ecological forecasting more dynamic and responsive, Dietze et al [-dietze2018] recently called for an increase in iterative near-term forecasting. Iterative near-term forecasting means making forecasts for the near future and making these forecasts repeatedly through a cycle of forecast evaluation, integration of updated data, and generation of new forecasts. This approach to forecasting has a number of advantages. Because forecasts are made 'near-term'---daily to annual forecasting instead of multi-decadal---predictions can be assessed more quickly and frequently, leading to more rapid model improvements [@tredennick2016; @dietze2018]. Because the forecasts are made repeatedly through time, new data can be integrated with each new iteration of the forecast cycle. This iterative approach to forecasting allows any changes in the state of the system that have occurred since the previous forecast to be incorporated and accounted for, thereby improving the accuracy of future forecasts [@dietze2018]. Iterative near-term forecasting has the potential to promote rapid improvement in the state of ecological forecasting by quickly identifying how models are failing, facilitating rapid testing of improved models, and incorporating updated data so models run with the most up-to-date information on the system available. While use of iterative near-term forecasting is often discussed in the context of policy and management [@clark2001; @Luo2011; @petchey2015], this approach to model testing can also be used to improve our basic understanding of ecological systems [@dietze2017]. For example, alternative mechanistic models can be competed against each other to see which model provides the best forecasts for near-term dynamics, thus providing insights into the relative importance of different processes driving dynamics of ecological systems [@dietze2018]. Whether deployed for basic or applied uses, iterative near-term forecasting incorporates a more dynamic interplay between models, predictions, and data that is clearly needed to improve ecological forecasting and our understanding of ecological systems more broadly.

Because iterative near-term forecasting requires a dynamic interplay of models, predictions, and data, Dietze et al [-dietze2018] highlight approaches to data management, model construction and evaluation, and cyberinfrastructure that are necessary to effectively implement this type of forecasting (Box 1). Data to be used for iterative near-term forecasting needs to be widely accessible, which requires data to be released quickly under open licenses [@dietze2018; @vargas2017] and structured so that it can be used easily by a variety of researchers and in multiple modeling approaches [@borer2009; @strasser2011]. Models need to be able to deal with uncertainty, in the predictors and the predictions, to properly convey uncertainty in the resulting forecasts [@diniz-filho2009]. Multiple models should be developed, both to assess which models are performing best [@dietze2018] and to facilitate combining models to form ensemble predictions which tend to perform better than single models [@araujo2007; @diniz-filho2009]. Ensuring that data and models are regularly updated and new forecasts are made requires cyberinfrastructure to automate data processing, model fitting, prediction, model evaluation, forecast visualization, and archiving. In combination, these approaches should allow forecasts to be easily rerun and evaluated as new data becomes available [Box 1; @dietze2018].

@@

In addition to automatically running software pipelines, the other key component of "continuous analysis" is making sure that the pipelines will continue to run even as software dependencies change [@beaulieu2017]. Many of us have experienced the frustrations that can occur when software updates (e.g., changes in R package versions) create errors in previously functional code. We experienced this issue when a package one of our models relies on, `tscount` [@liboschik2015], was temporarily removed from CRAN (the R package repository) and, therefore, could not be installed in the usual way. This broke our forecasting pipeline because we could no longer run models that used that package. To make our pipeline robust to changes in external software dependencies, we follow Beaulieu and Greene's [-@beaulieu2017] recommendation to use software containers. Software containers are standalone packages that contain copies of everything you need to run some piece of software, including the operating system. Once created, a software container is basically a time capsule, it contains all the software dependencies in the exact state used to develop and run the software. If those dependencies change (or disappear) in the wider world, they still exist, unchanged, in your container. We use an existing platform, Docker [@Merkel2014], to store an exact image of the complete software environment for running the forecasts. Docker also allows a specified set of packages to be used consistently across different computer and server environments. Using containers allows us to update to new package versions after testing and making any necessary changes to the data processing and analysis code. We use a container created by the Rocker project which is a Docker image with many important R packages (i.e. tidyverse) pre-installed  [@boettiger2017]. We add our code and dependencies to this existing Rocker image to create a software container for our forecasting pipeline. In combination, the automated running of the pipeline (continuous integration) and the guarantee it will not stop working unexpectedly due to software dependencies (via a software container) allows continuous analysis to serve as the glue that connects all stages of the forecasting pipeline.

@@

Following the recommendations of Dietze et al [-dietze2018], we developed an automated iterative forecasting system (Figure 1) to support repeated forecasting of an ecological system. Our forecasting system automatically acquires and processes the newest data, refits the models, makes new forecasts, publicly archives those forecasts, and presents both the current forecast and information on how previous forecasts performed. Every week our forecasting system generates a new set of forecasts with no human intervention, except for the entry of new field data. This ensures that forecasts based on the most recent data are always available and allows us to rapidly assess the performance of multiple forecasting models for a number of different states of the system, including the abundances of individual species and community-level variables such as total abundance. To create this iterative near-term forecasting system, we used R to process data and conduct analyses, and we leveraged already existing services (i.e. GitHub, Travis, Docker) for more complicated cyberinfrastructure tasks. Thus, our approach to developing iterative near-term forecasting infrastructure provides an example for how short-term ecological forecasting systems can be initially developed.

@@

A list of some of the key practices developed by Dietze et al [-dietze2018] for facilitating iterative near-term ecological forecasting and discussion of why these practices are important.

>>>>

Forecasting the future state of ecological systems is important for management, conservation, and evaluation of our basic understanding of how ecology operates [@clark2001; @tallis2006; @diaz2015; @dietze2017]. In 2001, Clark et al. [-@clark2001] called for a more central role of forecasting in ecology, and since then an increasing number of ecological forecasts are being published. However, most of these forecasts are made once, published, and never assessed or updated. Without assessment, we have limited information on how much confidence to place in our predictions; without regular updates, forecasts lack the most up-to-date information as conditions change [@dietze2018]. This lack of both regular assessment and active updating has limited the progress of ecological forecasting and hindered our ability to make useful and reliable predictions. More regular updating and assessment will advance ecological forecasting as a field by speeding up the identification of the best models for forecasting specific issues in specific ecosystems and contribute overall to our understanding of how to best design forecasting approaches for ecology. For ecological forecasting to mature as a field, we need to change how we produce and interact with forecasts, creating a more dynamic interplay between model development, prediction generation, and incorporation of new data and information [@dietze2018].

With the goal of making ecological forecasting more dynamic and responsive, Dietze et al [-@dietze2018] recently called for an increase in iterative near-term forecasting. Iterative near-term forecasting means making forecasts for the near future and making these forecasts repeatedly through a cycle of forecast evaluation, integration of updated data, and generation of new forecasts. This approach to forecasting has a number of advantages. Because forecasts are made 'near-term'---daily to annual forecasting instead of multi-decadal---predictions can be assessed more quickly and frequently, leading to more rapid model improvements [@tredennick2016; @dietze2018]. Because the forecasts are made repeatedly through time, new data can be integrated with each new iteration of the forecast cycle. This iterative approach to forecasting allows any changes in the state of the system that have occurred since the previous forecast to be incorporated and accounted for, thereby improving the accuracy of future forecasts [@dietze2018]. Iterative near-term forecasting has the potential to promote rapid improvement in the state of ecological forecasting by quickly identifying how models are failing, facilitating rapid testing of improved models, and incorporating updated data so models run with the most up-to-date information on the system available. While use of iterative near-term forecasting is often discussed in the context of policy and management [@clark2001; @luo2011; @petchey2015], this approach to model testing can also be used to improve our basic understanding of ecological systems [@dietze2017]. For example, alternative mechanistic models can be competed against each other to see which model provides the best forecasts for near-term dynamics, thus providing insights into the relative importance of different processes driving dynamics of ecological systems [@dietze2018]. Whether deployed for basic or applied uses, iterative near-term forecasting incorporates a more dynamic interplay between models, predictions, and data that is clearly needed to improve ecological forecasting and our understanding of ecological systems more broadly.

Because iterative near-term forecasting requires a dynamic interplay of models, predictions, and data, Dietze et al [-@dietze2018] highlight approaches to data management, model construction and evaluation, and cyberinfrastructure that are necessary to effectively implement this type of forecasting (Box 1). Data to be used for iterative near-term forecasting needs to be widely accessible, which requires data to be released quickly under open licenses [@dietze2018; @vargas2017] and structured so that it can be used easily by a variety of researchers and in multiple modeling approaches [@borer2009; @strasser2011]. Models need to be able to deal with uncertainty, in the predictors and the predictions, to properly convey uncertainty in the resulting forecasts [@diniz-filho2009]. Multiple models should be developed, both to assess which models are performing best [@dietze2018] and to facilitate combining models to form ensemble predictions which tend to perform better than single models [@araujo2007; @diniz-filho2009]. Ensuring that data and models are regularly updated and new forecasts are made requires cyberinfrastructure to automate data processing, model fitting, prediction, model evaluation, forecast visualization, and archiving. In combination, these approaches should allow forecasts to be easily rerun and evaluated as new data becomes available [Box 1; @dietze2018].

@@

In addition to automatically running software pipelines, the other key component of "continuous analysis" is making sure that the pipelines will continue to run even as software dependencies change [@beaulieu2017]. Many of us have experienced the frustrations that can occur when software updates (e.g., changes in R package versions) create errors in previously functional code. We experienced this issue when a package one of our models relies on, `tscount` [@liboschik2015], was temporarily removed from CRAN (the R package repository) and, therefore, could not be installed in the usual way. This broke our forecasting pipeline because we could no longer run models that used that package. To make our pipeline robust to changes in external software dependencies, we follow Beaulieu and Greene's [-@beaulieu2017] recommendation to use software containers. Software containers are standalone packages that contain copies of everything you need to run some piece of software, including the operating system. Once created, a software container is basically a time capsule, it contains all the software dependencies in the exact state used to develop and run the software. If those dependencies change (or disappear) in the wider world, they still exist, unchanged, in your container. We use an existing platform, Docker [@merkel2014], to store an exact image of the complete software environment for running the forecasts. Docker also allows a specified set of packages to be used consistently across different computer and server environments. Using containers allows us to update to new package versions after testing and making any necessary changes to the data processing and analysis code. We use a container created by the Rocker project which is a Docker image with many important R packages (i.e. tidyverse) pre-installed  [@boettiger2017]. We add our code and dependencies to this existing Rocker image to create a software container for our forecasting pipeline. In combination, the automated running of the pipeline (continuous integration) and the guarantee it will not stop working unexpectedly due to software dependencies (via a software container) allows continuous analysis to serve as the glue that connects all stages of the forecasting pipeline.

@@

Following the recommendations of Dietze et al [-@dietze2018], we developed an automated iterative forecasting system (Figure 1) to support repeated forecasting of an ecological system. Our forecasting system automatically acquires and processes the newest data, refits the models, makes new forecasts, publicly archives those forecasts, and presents both the current forecast and information on how previous forecasts performed. Every week our forecasting system generates a new set of forecasts with no human intervention, except for the entry of new field data. This ensures that forecasts based on the most recent data are always available and allows us to rapidly assess the performance of multiple forecasting models for a number of different states of the system, including the abundances of individual species and community-level variables such as total abundance. To create this iterative near-term forecasting system, we used R to process data and conduct analyses, and we leveraged already existing services (i.e. GitHub, Travis, Docker) for more complicated cyberinfrastructure tasks. Thus, our approach to developing iterative near-term forecasting infrastructure provides an example for how short-term ecological forecasting systems can be initially developed.

@@

A list of some of the key practices developed by Dietze et al [-@dietze2018] for facilitating iterative near-term ecological forecasting and discussion of why these practices are important.
